% chapters/chapter2.tex
% Chapter 2: Quantum Clustering - Main file that includes algorithm subsections

\chapter{Quantum Clustering: A New Frontier in Image Segmentation}

\section{Introduction: Bridging Quantum Mechanics and Computer Vision}

The intersection of quantum computing and computer vision represents an emerging field with significant potential for advancing image segmentation capabilities~\cite{wang2022review,cai2018survey,aburaed2017advances}. This chapter explores how quantum mechanical principles can be leveraged to develop more efficient clustering algorithms for image analysis~\cite{ruan2021qip,chakraborty2022challenges}.

Quantum computing fundamentally differs from classical computing in its ability to exploit quantum mechanical phenomena—superposition, entanglement, and interference—to process information~\cite{sood2024quantum}. These properties enable quantum algorithms to explore solution spaces in ways that are impossible for classical computers~\cite{li2020quantum,ramezani2020mlquantum}.

\begin{keyconceptbox}[Quantum Computing for Clustering]
The key insight driving quantum clustering research is that many clustering problems can be reformulated as~\cite{aimeur2007clustering,xu2015clustering}:
\begin{itemize}
    \item Distance estimation problems (quantum advantage via amplitude encoding)
    \item Eigenvalue problems (quantum advantage via quantum phase estimation)
    \item Optimization problems (quantum advantage via QAOA or quantum annealing)
\end{itemize}
\end{keyconceptbox}

\subsection{The NISQ Era and Its Implications}

Current quantum computers operate in the Noisy Intermediate-Scale Quantum (NISQ) era, characterized by:

\begin{itemize}
    \item \textbf{Limited qubit counts:} Typically 50-1000 qubits
    \item \textbf{High error rates:} Gate errors of $10^{-3}$ to $10^{-2}$
    \item \textbf{Limited coherence times:} Microseconds to milliseconds
    \item \textbf{Restricted connectivity:} Not all qubits can interact directly
\end{itemize}

\begin{warningbox}
These limitations mean that theoretically optimal quantum algorithms may not be practical on current hardware. Much research focuses on variational and hybrid classical-quantum approaches that are more noise-tolerant.
\end{warningbox}

\subsection{From Classical to Quantum: Key Transformations}

Applying quantum computing to image segmentation requires several key transformations:

\begin{enumerate}
    \item \textbf{Data encoding:} Converting classical image data into quantum states
    \item \textbf{Algorithm design:} Developing quantum circuits that perform clustering operations
    \item \textbf{Measurement and interpretation:} Extracting classical cluster assignments from quantum measurements
\end{enumerate}

\section{Approaches to Quantum Image Representation and Processing}

A fundamental challenge in quantum image processing is encoding classical image data into quantum states~\cite{yan2014qip}. Several representations have been proposed, each with distinct advantages for different applications~\cite{cai2018survey,wang2022review}.

\subsection{Flexible Representation of Quantum Images (FRQI)}

FRQI encodes a $2^n \times 2^n$ grayscale image using $2n + 1$ qubits:

\begin{definitionbox}[FRQI Representation]
\begin{equation}
    \ket{\text{FRQI}} = \frac{1}{2^n} \sum_{i=0}^{2^{2n}-1} \left(\cos\theta_i\ket{0} + \sin\theta_i\ket{1}\right) \otimes \ket{i}
\end{equation}
where $\theta_i \in [0, \pi/2]$ encodes the grayscale value of pixel $i$, and $\ket{i}$ encodes the pixel position.
\end{definitionbox}

\textbf{Advantages:}
\begin{itemize}
    \item Compact representation: $2n + 1$ qubits for $2^{2n}$ pixels
    \item Natural encoding for grayscale images
\end{itemize}

\textbf{Limitations:}
\begin{itemize}
    \item Complex state preparation requiring $O(2^{2n})$ gates
    \item Limited grayscale precision (single qubit for intensity)
\end{itemize}

\subsection{Novel Enhanced Quantum Representation (NEQR)}

NEQR improves upon FRQI by using a basis encoding for pixel intensities:

\begin{definitionbox}[NEQR Representation]
\begin{equation}
    \ket{\text{NEQR}} = \frac{1}{2^n} \sum_{y=0}^{2^n-1} \sum_{x=0}^{2^n-1} \ket{f(y,x)}\ket{yx}
\end{equation}
where $\ket{f(y,x)} = \ket{c_0^{yx}c_1^{yx}\ldots c_{q-1}^{yx}}$ is the $q$-bit binary representation of the grayscale value at position $(y,x)$.
\end{definitionbox}

\textbf{Advantages:}
\begin{itemize}
    \item Higher precision: $q$ bits for intensity values (e.g., $q=8$ for 256 levels)
    \item Simpler image operations through bit-level manipulations
    \item More efficient for certain quantum image processing operations
\end{itemize}

\subsection{Amplitude Encoding for Feature Vectors}

For clustering applications, amplitude encoding is particularly important as it allows efficient representation of high-dimensional feature vectors:

\begin{definitionbox}[Amplitude Encoding]
A normalized vector $\mathbf{x} = (x_0, x_1, \ldots, x_{N-1})^T$ with $\|\mathbf{x}\| = 1$ can be encoded as:
\begin{equation}
    \ket{\mathbf{x}} = \sum_{i=0}^{N-1} x_i \ket{i}
\end{equation}
requiring only $\lceil\log_2 N\rceil$ qubits to represent $N$ amplitudes.
\end{definitionbox}

\begin{tipbox}
Amplitude encoding enables exponential compression: a feature vector with $N = 2^n$ components requires only $n$ qubits. This is crucial for encoding high-dimensional image features efficiently.
\end{tipbox}

\subsection{Quantum Distance Estimation}

A key operation for clustering is computing distances between data points. Quantum computers can estimate distances using the swap test:

\begin{equation}
    |\braket{\mathbf{x}|\mathbf{y}}|^2 = 1 - \frac{d(\mathbf{x}, \mathbf{y})^2}{2}
\end{equation}

where $d(\mathbf{x}, \mathbf{y})$ is the Euclidean distance between normalized vectors.

The swap test circuit measures the overlap between two quantum states with $O(1)$ quantum operations (after state preparation), compared to $O(d)$ classical operations for $d$-dimensional vectors.

\section{A Comprehensive Review of Quantum Segmentation Algorithms}

This section presents a detailed analysis of quantum clustering algorithms that have been proposed for image segmentation. Rather than an exhaustive catalog, we focus on representative methods that illustrate the diverse strategies for leveraging quantum computation in clustering tasks.

\subsection{Organization and Scope}

The algorithms reviewed here are organized according to their underlying \textbf{quantum computational paradigm}, progressing from hardware-native approaches to hybrid and quantum-inspired methods:

\begin{enumerate}
    \item \textbf{Quantum Annealing Methods} (Section~\ref{sec:quantum-annealing}): We begin with Q-Seg and related approaches that exploit quantum annealing hardware (e.g., D-Wave systems). These methods reformulate segmentation as Quadratic Unconstrained Binary Optimization (QUBO) problems, leveraging quantum tunneling to escape local minima. They represent the most direct path to real quantum hardware execution today.
    
    \item \textbf{Gate-Model QAOA Methods} (Sections~\ref{sec:quantum-logismos} and \ref{sec:qcffcm}): Next, we examine algorithms based on the Quantum Approximate Optimization Algorithm (QAOA), which operates on gate-based quantum computers. QuantumLOGISMOS applies QAOA to geometric-constrained medical image segmentation, while QCFFCM integrates QAOA into fuzzy clustering for high-resolution scientific imagery.
    
    \item \textbf{Quantum-Inspired Classical Methods} (Section~\ref{sec:ffqoak}): Finally, we present FFQOAK, a quantum-inspired approach that simulates quantum dynamics on classical hardware. This represents an important category: methods that capture quantum algorithmic ideas without requiring quantum hardware.
\end{enumerate}

\begin{keyconceptbox}[Why These Algorithms?]
The selected algorithms illustrate three fundamental questions in quantum image segmentation:
\begin{itemize}
    \item \textbf{Problem formulation:} How is segmentation encoded as a quantum-compatible optimization? (QUBO, Ising Hamiltonians, variational objectives)
    \item \textbf{Hardware requirements:} What quantum resources are needed? (Annealers vs. gate-model vs. classical simulation)
    \item \textbf{Scalability trade-offs:} How do theoretical advantages translate to practical image sizes?
\end{itemize}
\end{keyconceptbox}

\subsection{Distinguishing Characteristics}

To help navigate the diversity of approaches, we highlight key differentiators:

\begin{table}[H]
\centering
\small
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Algorithm} & \textbf{Paradigm} & \textbf{Classical Foundation} & \textbf{Primary Innovation} \\
\midrule
Q-Seg & Quantum Annealing & Graph cuts & QUBO min-cut formulation \\
QuantumLOGISMOS & QAOA (Gate-model) & LOGISMOS & Multiple optimal solutions \\
QCFFCM & QAOA (Gate-model) & Fuzzy C-Means & 3-ADMM-H hybrid framework \\
FFQOAK & Quantum-inspired & K-Means & Quantum dynamics simulation \\
\bottomrule
\end{tabular}
\caption{Overview of quantum segmentation algorithms covered in this review}
\label{tab:algo-overview}
\end{table}

Each algorithm builds upon a well-established classical segmentation technique (graph cuts, LOGISMOS, FCM, or K-means) and introduces quantum enhancements at specific computational bottlenecks. This hybrid philosophy---quantum acceleration of classical frameworks---reflects the practical reality of the NISQ era.

\begin{notebox}[Reading Guide]
Readers interested in \textbf{immediate quantum hardware deployment} should focus on Q-Seg (quantum annealing). Those exploring \textbf{medical imaging applications} will find QuantumLOGISMOS most relevant. For \textbf{large-scale practical segmentation} without quantum hardware access, FFQOAK and QCFFCM offer deployable solutions today.
\end{notebox}

The section concludes with a \textbf{comparative analysis} that synthesizes these approaches, providing guidance on algorithm selection based on application requirements, available hardware, and performance priorities.

%% Include individual algorithm files
%% Each algorithm is in its own file for parallel editing

\input{chapters/algorithms/quantum-annealing}
\input{chapters/algorithms/quantum-logismos}
\input{chapters/algorithms/quantum-ffqoak}
\input{chapters/algorithms/quantum-fuzzy-qcffcm}
\input{chapters/algorithms/comparative-analysis}

%% ==============================================
%% HOW TO ADD A NEW ALGORITHM:
%% ==============================================
%% 1. Create a new file in chapters/algorithms/ (e.g., quantum-newmethod.tex)
%% 2. Start the file with \subsection{Your Algorithm Name}
%% 3. Add your content (subsubsections, equations, boxes, etc.)
%% 4. Add an \input line above this comment block:
%%    \input{chapters/algorithms/quantum-newmethod}
%% 5. Update the comparative analysis table in comparative-analysis.tex
%% ==============================================
