% chapters/conventions.tex
% Conventions Used in This Report

\chapter*{Conventions Used in This Report}
\addcontentsline{toc}{chapter}{Conventions Used in This Report}
\thispagestyle{plain}

This report uses the following typographical conventions and special callout boxes to highlight important information:

\vspace{0.8cm}

\begin{notebox}[Note]
A \textbf{Note} box (blue) contains useful background information, additional context, or interesting observations that supplement the main text. Notes provide helpful details that are relevant but not critical to understanding the core material.
\end{notebox}

\vspace{0.5cm}

\begin{tipbox}[Tip]
A \textbf{Tip} box (teal) offers practical advice, best practices, or suggestions that can help you better understand or apply the concepts being discussed. Tips often highlight shortcuts or insights gained from experience.
\end{tipbox}

\vspace{0.5cm}

\begin{warningbox}[Warning]
A \textbf{Warning} box (orange) indicates important caveats, potential pitfalls, or common mistakes to avoid. Pay attention to warnings to prevent misunderstandings or errors in applying the discussed methods.
\end{warningbox}

\vspace{0.5cm}

\begin{cautionbox}[Caution]
A \textbf{Caution} box (red) signals critical information that could lead to serious problems if ignored. Cautions highlight significant limitations, dangerous assumptions, or fundamental constraints.
\end{cautionbox}

\vspace{0.5cm}

\begin{definitionbox}[Definition]
A \textbf{Definition} box (gray) provides formal definitions of key terms, mathematical concepts, or technical vocabulary. These boxes establish precise meanings for important terminology used throughout the report.
\end{definitionbox}

\vspace{0.5cm}

\begin{keyconceptbox}[Key Concept]
A \textbf{Key Concept} box (orange accent) highlights fundamental ideas, core principles, or essential takeaways that are central to understanding the material. These boxes summarize the most important points.
\end{keyconceptbox}

\vspace{0.5cm}

\begin{examplebox}[Example]
An \textbf{Example} box (light gray) presents concrete illustrations, worked problems, or practical demonstrations of the concepts being explained. Examples help bridge theory and application.
\end{examplebox}

\vspace{0.8cm}

\noindent\textbf{Additional conventions:}
\begin{itemize}
    \item \textit{Italics} are used for emphasis and for introducing new terms.
    \item \textbf{Bold} is used for important concepts and terminology.
    \item \texttt{Monospace font} is used for code, algorithms, and technical notation.
    \item Mathematical expressions use standard notation, with quantum states written in Dirac notation (e.g., $\ket{\psi}$, $\bra{\phi}$, $\braket{\phi|\psi}$).
\end{itemize}
